On the adhesivity properties of equivalence graphs\thanks{???????}

### 2025-02-03 icgt/main.tex
cose

\spnewtheorem*{notation}{Notation}{\bfseries}{}

@@

\begin{theorem}
	\label{thm:slice-functors}
	Let $\X$ be an $\mathcal{M}$-adhesive category, then the following hold true:
	\begin{enumerate}
		\item if $\Y$ is an $\mathcal{N}$-adhesive category $L\colon \Y\to \A$ a functor preserving $\mathcal{N}$-pushouts and $R\colon \X\to A$ one preserving pullbacks, then $\comma{L}{R}$ is $\comma{\mathcal{N}}{\mathcal{M}}$-adhesive, where
		\[\comma{\mathcal{N}}{\mathcal{M}}:=\{(h,k) \in \mor(\comma{L}{R}) \mid h\in \mathcal{N}, k\in \mathcal{M}\}\]
		\item for every object $X$
		the categories $\X/X$  and $X/X$ are, respectively, $\mathcal{M}/X$-adhesive and $X/\mathcal{M}$-adhesive, where
		\[\mathcal{M}/X:=\{m\in \mor(\X/X) \mid m\in
		\mathcal{M}\} \hspace{5pt} X/\mathcal{M}:=\{m\in \mor(X/\X) \mid m\in \mathcal{M}\}\]
		\item for every small category $\Y$, the category $\X^\Y$ of
		functors $\Y\to \X$ is $\mathcal{M}^{\Y}$-adhesive, where
		$\mathcal{M}^{\Y}:=\{\eta \in \mor(\X^\Y) \mid \eta_Y \in
		\mathcal{M} \text{ for every } Y\in \Y\}$;
		\item given a  full and faithful $I\colon \Y\to \X$ and a class $\mathcal{M}'$ of arrows in $\Y$ stable under pullbacks, pushout, containing all isomorphism and closed under composition an decomposition, if image of $\mathcal{M}'$ under $I$ is contained in $\mathcal{M}$ and $a$
	\end{enumerate} 
\end{theorem}

@@

A pivotal role will be played by the Kleene star $(-)^\star\colon \Set\to \Set$ sending a set to the free monoid on it \cite{...}. In particular, given a set $X$, $X^\star$ is the set of functions with a natural number as domain. Then the neutral element is $?_X\colon \emptyset \to X$ and the composition of $f\colon n\to X$ with $g\colon m\to X$ is simply the arrow $n+m\to X$ induced by them. If $h\colon X\to Y$ is an arrow in $\Set$, then $h^\star \colon X^\star\to Y^\star$ is the composition with $h$.

\subsection{The category of hypergraphs}

We will start this section with the definition of hypergraphs and we will see how to label them with an algebraic signature.

\begin{definition}A \emph{hypergraph} is a 4-uple $\mathcal{G}:=(E_\mathcal{G}, V_\mathcal{G}, s_\mathcal{G}, t_\mathcal{G})$ made by two sets $E_\mathcal{G}$ and $V_\mathcal{G}$, whose elements are called respectively \emph{hyperedges} and \emph{nodes}, plus a pair of \emph{source} and \emph{target} functions  $s_\mathcal{G}, t_\mathcal{G}\colon E_\mathcal{G}\rightrightarrows V_\mathcal{G}^\star$. A \emph{hypergraph morphism} $(E_\mathcal{G}, V_\mathcal{G}, s_\mathcal{G}, t_\mathcal{G})\to (E_\mathcal{H}, V_\mathcal{H}, s_\mathcal{H}, t_\mathcal{H})$ is a pair $(h,k)$ of functions $h\colon E_\mathcal{G}\to E_\mathcal{H}$, $k\colon V_\mathcal{G}\to V_\mathcal{H}$ such that the following diagrams commute.

	\[\xymatrix{ E_{\mathcal{G}} \ar[d]_{h} \ar[r]^{s_{\mathcal{G}}}& V^\star_{\mathcal{G}}  \ar[d]^{k^\star}& E_{\mathcal{G}} \ar[r]^{t_{\mathcal{G}}} \ar[d]_{h} & V^\star_{\mathcal{G}} \ar[d]^{k^\star}  \\ E_{\mathcal{G}} \ar[r]_{s_{\mathcal{H}}} & V^\star_{\mathcal{H}} & E_{\mathcal{G}} \ar[r]_{t_{\mathcal{H}}} & V^\star_{\mathcal{H}} }\]
	We define $\hyp$ to be the resulting category.
\end{definition}

Let $\pro^\star$ be the functor sending $X$ to $X^\star\times X^\star$, then we can present $\hyp$ as a comma category.
\begin{proposition}\label{prop:com}
	$\hyp$ is isomorphic to $\comma{\id{\Set}}{\pro^\star}$
\end{proposition}
\begin{proof}
	 Define two functors $F\colon  \hyp \to \comma{\id{\Set}}{\pro^\star} $
	and $G\colon \comma{\id{\Set}}{\pro^\star}\to \hyp$	 as follows 
	\begin{align*}
		\functor[l]{(E_\mathcal{{G}}, V_\mathcal{{G}}, s_\mathcal{{G}}, t_\mathcal{{G}})}{(f,g)}{\left(E_\mathcal{{H}}, V_\mathcal{{H}}, s_\mathcal{{H}}, t_\mathcal{{H}}\right)}
		\functormapsto
		\rfunctor{\left(E_\mathcal{{G}}, V_\mathcal{{G}}, 	\left(s_\mathcal{{G}}, t_\mathcal{{G}}\right)\right)}{(f, g) }{\left(E_\mathcal{{H}}, V_\mathcal{{H}}, \left(s_\mathcal{{H}}, t_\mathcal{{H}}\right)\right)}
	\end{align*}
	 \begin{align*}
		\functor[l]{(E_\mathcal{{G}}, V_\mathcal{{G}}, p_{\mathcal{G}})}{(f,g)}{(E_\mathcal{{H}}, V_\mathcal{{H}}, p_{\mathcal{H}})}
		\functormapsto
		\rfunctor{(E_\mathcal{{G}}, V_\mathcal{{G}}, 	\pi_1\circ  p_{\mathcal{G}}, \pi_2\circ  p_{\mathcal{G}})}{(f, g) }{(E_\mathcal{{H}}, V_\mathcal{{H}}, 	\pi_1\circ  p_{\mathcal{H}}, \pi_2\circ  p_{\mathcal{H}})}
	\end{align*} 
	Now it is immediate to notice that they are one the inverse of the other.
\end{proof}

We can show that the  Kleene star preserves pullbacks (see also \cite[Sec. 3]{carboni1995connected}  and \cite[Ch.4]{leinster2004higher} for details and a deeper and more conceptual approach).\index{Kleene star}

\begin{proposition}\label{prop:pb}
	The functor $(-)^\star$ preserves pullbacks.
\end{proposition}
\begin{proof}
	Suppose that a pullbacks square as the one below is given.
	\[\xymatrix{P \ar[r]^{p_1} \ar[d]_{p_2}& X \ar[d]^{f}\\ Y \ar[r]_{g} & Z}\]
	and consider the solid part of the following diagram.
	\[\xymatrix{Z\ar@{.>}[dr]^{t} \ar@/^.3cm/[drr]^{t_1} \ar@/_.3cm/[ddr]_{t_2}\\&P^\star \ar[r]^{p^\star_1} \ar[d]_{p^\star_2}& X^\star \ar[d]^{f^\star}\\ &Y^\star \ar[r]_{g^\star} & Z^\star}\]
	For every $z\in Z$ we have $t_1(z)\colon n\to X$ and $t_2(z)\colon m\to Y$ such that
	\[f^\star(t_1(x))=g^\star(t_2(z))\]
	In particular this entails that $n=m$ and that there is $t(z)\colon n\to P$ as in the diagram below 
	\[\xymatrix{n\ar[dr]^{t(z)} \ar@/^.3cm/[drr]^{t_1(z)} \ar@/_.3cm/[ddr]_{t_2(z)}\\&P \ar[r]^{p_1} \ar[d]_{p_2}& X \ar[d]^{f}\\ &Y \ar[r]_{g} & Z}\]
	But this is equivalent to say that the dotted $t\colon Z\to P$ exists, while its uniqueness follows at once from the universal property of the pullback with which we started.
\end{proof}

@@

\begin{corollary}\label{prop:hypadh}
	$\hyp$ is an adhesive category.
\end{corollary}
\begin{proof}
	$(-)^\star$ preserves pullbacks by \Cref{prop:pb}, while $\pro$ is continuous by definition, thus the thesis follows from This follows from \Cref{comma,prop:com}.
\end{proof}

@@

\begin{proposition}\label{cor:left}
	The forgetful functor $U_{\catname{Hyp}}\colon \hyp \to \catname{Set}$ which sends an hypergraph $\mathcal{G}$ to its set of nodes has a left adjoint $\Delta_{\hyp}$.
\end{proposition}

\begin{remark}Since the initial object of $\catname{Set}$ is the empty set,  $\Delta_{\hyp}(X)$ is the hypergraph which has $X$ as set of nodes and $\emptyset$ as set of hyperedges and $?_X$ as both source and target function.
\end{remark}

Hypergraphs,  can be represented graphically. We will use dots to denote nodes and squares to denote hyperedges, the name of a node or of an hyperedge will be put near the corresponding dot or square. Sources and targets are represented by lines between dots and squares: the lines from the sources of an hyperedge will have an arrowhead in the middle pointing towards the hyperedge, while the lines to the targets will have arrowheads pointing to the target nodes.  We will decorate the arrow corresponding to the $i^{th}$ letter (i.e.~its value at $i-1$) of a target or a source with a label $i$.

@@

\subsubsection{$\hyp$ as a topos of presheaves}

By \Cref{lim} we already know that $\hyp$ has all pullbacks and by \Cref{prop:hypadh} we know that it is adhesive. Actually more can be proved about it: we can realize $\hyp$ as a topos of presheaves \cite{bonchi2022string}.

@@

Now, given $F\colon \catname{H}\to \catname{Set}$ we can define
\[E_F:=\sum_{k,l\in \mathbb{N}}F(k,l)\]
For every element $x$ of $F(k,l)$ we can put
\begin{align*}
	s^F_{k,l}(x)\colon k\to F(\bullet) \quad i \mapsto F(f_i)(x)\qquad 
	t^F_{k,l}(x)\colon l\to F(\bullet) \quad i \mapsto F(f_{i+k})(x)
\end{align*}
obtaining
$s_F, t_F\colon E_F\rightrightarrows \mathcal{F}(\bullet)^{\star}$. Let $\mathcal{G}_F$ be the resulting hypergraph. Now, every $\eta\colon F\rightarrow H$ in $\catname{Set}^{\catname{H}}$ has components $\eta_{k,l}\colon F(k,l)\to H(k,l)$, $\eta_{\bullet}\colon F(\bullet)\to H(\bullet)$, thus it induces a function $\hat{\eta}\colon E_F\rightarrow E_H$ such that the following squares commute
\[\xymatrix{ E_F \ar[r]^{s_F} \ar[d]_{\hat{\eta}}& F(\bullet)^\star \ar[d]^{\eta^\star_\bullet} & E_F \ar[r]^{t_F} \ar[d]_{\hat{\eta}} & F(\bullet)^\star \ar[d]^{\eta^\star_\bullet}\\ E_H \ar[r]_{s_H} & H(\bullet)^\star & E_H \ar[r]_{t_H}& H(\bullet)^\star}\]

This is equivalent to say that $\eta$ induces a morphism $(\hat{\eta}, \eta_{\bullet})\colon \mathcal{G}_F\to \mathcal{G}_H$. It is now clear that sending $F$ to $\mathcal{G}_F$ and $\eta$ to $(\hat{\eta}, \eta_{\bullet})$ defines a faithful functor $\mathcal{G}_{-}\colon \catname{Set}^{\catname{H}}\to \hyp$.

\begin{proposition}
	$\hyp$ is equivalent to the category $\catname{Set}^{\catname{H}}$.
\end{proposition}
\begin{proof}
	Let $X$ be a set, for every $n\in \mathbb{N}$ define
	\[X_{n}:= \left\{w\in X^{\star} \mid \dom(w)=n\right\}\]
	In particular, if $F\colon \catname{H}\to \catname{Set}$ then the image of the coprojection $\iota^{F}_{k,l}\colon F(k,l)\to E_F$ is the intersection
	\[s^{-1}_F\left(F(\bullet)_{k}\right)\cap t^{-1}_F\left(F(\bullet)_{l}\right) \]

	We are now ready to that $\mathcal{G}_{-}$ is full and essentially surjective.
	\begin{itemize}
		\item For fullness, let $(f,g)\colon \mathcal{G}_F\to \mathcal{G}_{H}$ be a morphism of hypergraphs and define $f_{k,l}$ to be $f\circ \iota^F_{k,l}$, the composition of $h$ with Now, if $x\in F(k,l)$ then 
		\[\begin{split}
			s_H (f_{k,l} (x))&=s_H \left(f \left(\iota^F_{k,l} (x)\right)\right)\\&=g^{\star}\hspace{-1pt}\left(s_{F} \hspace{-1pt}\left(\iota^F_{k,l} (x)\right)\right)
		\end{split}\qquad \begin{split}
			t_H (f_{k,l} (x))&=s_t\hspace{-1pt} \left(f \hspace{-1pt}\left(\iota^F_{k,l} (x)\right)\right)\\&=g^{\star}\hspace{-1pt} \left(t_{F}\hspace{-1pt} \left(\iota^F_{k,l} (x)\right)\right)
		\end{split}\]
		Therefore there exists $\eta_{k,l}\colon F(k,l)\to H(k,l)$ fitting in the diagram below
		\[\xymatrix{F(k,l) \ar[r]^-{\iota^{F}_{k,l}} \ar@{.>}[d]_{\eta_{k,l}}& E_F \ar[d]^{f_{k,l}}\\ H(k,l) \ar[r]_-{\iota^{H}_{k,l}}& E_H}\]
		Define $\eta_{\bullet}\colon F(\bullet)\to H(\bullet)$ simply as $g^\star$, then  the collection of all the $\eta_{k,l}$ and of $\eta_\bullet$ defines a natural transformation $\eta\colon F\to H$. Indeed, if $f_i\colon (k,l)\to \bullet$ we have:
		\[\xymatrix{F(k,l) \ar@/^.4cm/[rr]^{s^F_{k,l}} \ar[r]_-{\iota^{F}_{k,l}} \ar[d]_{\eta_{k,l}}& E_F \ar[d]_{f_{k,l}} \ar[r]_-{s_F}& F(\bullet)^\star \ar[d]^{g^\star}&F(k,l) \ar@/^.4cm/[rr]^{t^F_{k,l}}\ar[r]_-{\iota^{F}_{k,l}} \ar[d]_{\eta_{k,l}}& E_F \ar[d]_{f_{k,l}} \ar[r]_-{t_F}& F(\bullet)^\star \ar[d]^{g^\star}\\ H(k,l) \ar@/_.4cm/[rr]_{s^H_{k,l}} \ar[r]^-{\iota^{H}_{k,l}}& E_H \ar[r]^-{s_H}& H(\bullet)^\star& H(k,l) \ar@/_.4cm/[rr]_{t^H_{k,l}} \ar[r]^-{\iota^{H}_{k,l}}& E_H \ar[r]^-{t_H}& H(\bullet)^\star}\]
		Thus if $i<k$ then
		\begin{align*}
			\eta_\bullet(F(f_i)(x))&=g(F(f_i)(x))\\&=g\hspace{-1pt}\left(s^F_{k,l}(x)(i)\right)\\&=g^\star\hspace{-1pt}\left(s^F_{k,l}(x)\right)(i)\\&=s^H_{k,l}\hspace{-1pt}\left(\eta_{k,l}(x)\right)(i)\\&=F(f_i)(\eta_{k,l}(x))
		\end{align*}
		while, if $k\leq i < k+l-1$
		\begin{align*}
			\eta_\bullet(F(f_i)(x))&=g(F(f_i)(x))\\&=g\hspace{-1pt}\left(t^F_{k,l}(x)(i)\right)\\&=g^\star\hspace{-1pt}\left(t^F_{k,l}(x)\right)(i)\\&=t^H_{k,l}\hspace{-1pt}\left(\eta_{k,l}(x)\right)(i)\\&=F(f_i)(\eta_{k,l}(x))
		\end{align*}
		Finally, by contruction it is clear that $(\hat{\eta}, \eta_{\bullet})=(f,g)$. 
		\item Given an hypergraph $\mathcal{G}=(E_\mathcal{G}, V_\mathcal{G}, s_\mathcal{G}, t_\mathcal{G})$ we can define 
		\[F_{\mathcal{G}}(k,l):=s_\mathcal{G}^{-1}(V_k)\cap t_\mathcal{G}^{-1}(V_l) \qquad F_{\mathcal{G}}(\bullet):=V_\mathcal{G}\]
		Given $f_i\colon (k,l)\to \bullet$ we put
		\[F_{\mathcal{G}}(f_i)\colon F_{\mathcal{G}}(k,l)\to F_{\mathcal{G}}(\bullet) \qquad x\mapsto \begin{cases}
			s_\mathcal{G}(x)(i) & i<k\\
			t_\mathcal{G}(x) (i-k) &i\leq k < k+l-1
		\end{cases}  \]

		$F_{\mathcal{G}}$ so defined is a functor $\catname{H}\to \catname{Set}$ and for every $h\in E_\mathcal{G}$ there exists a unique pair $(k,l)$ such that $h\in F_{\mathcal{G}}(k,l) $, namely the pair $(\dom(s_\mathcal{G})(h), \dom(t_\mathcal{G})(h))$thus
		\[\sum_{k,l\in \mathbb{N}}F_{\mathcal{G}}(k,l)\simeq E\]
		Moreover, by construction $s_{F_{\mathcal{G}}}=s$ and $t_{F_{\mathcal{G}}}=t$, from which the thesis follows. 
	\end{itemize}
\end{proof}
As a corollary we get immediately the following.
\begin{corollary}
	$\hyp$ is a complete category.
\end{corollary}

>>>>

\spnewtheorem*{notation}{Notation}{\bfseries}{}
\spnewtheorem*{convention}{Convention}{\bfseries}{}

@@

\begin{theorem}
	\label{thm:slice-functors}
	Let $\X$ be an (strict) $\mathcal{M}$-adhesive category, then the following hold true:
	\begin{enumerate}
		\item if $\Y$ is an (strict) $\mathcal{N}$-adhesive category $L\colon \Y\to \A$ a functor preserving $\mathcal{N}$-pushouts and $R\colon \X\to A$ one preserving pullbacks, then $\comma{L}{R}$ is (strictly) $\comma{\mathcal{N}}{\mathcal{M}}$-adhesive, where
		\[\comma{\mathcal{N}}{\mathcal{M}}:=\{(h,k) \in \mor(\comma{L}{R}) \mid h\in \mathcal{N}, k\in \mathcal{M}\}\]
		\item for every object $X$
		the categories $\X/X$  and $X/X$ are, respectively, (strictly) $\mathcal{M}/X$-adhesive and (strictly) $X/\mathcal{M}$-adhesive, where
		\[\mathcal{M}/X:=\{m\in \mor(\X/X) \mid m\in
		\mathcal{M}\} \hspace{5pt} X/\mathcal{M}:=\{m\in \mor(X/\X) \mid m\in \mathcal{M}\}\]
		\item for every small category $\Y$, the category $\X^\Y$ of
		functors $\Y\to \X$ is (strictly) $\mathcal{M}^{\Y}$-adhesive, where
		$\mathcal{M}^{\Y}:=\{\eta \in \mor(\X^\Y) \mid \eta_Y \in
		\mathcal{M} \text{ for every } Y\in \Y\}$;
		\item if $\Y$ is a full subcategory of $\X$ closed in it under pullbacks and $\mathcal{M}$-pushouts, then $\Y$ is (strictly) $\mathcal{N}$-adhesive for every class of arrows $\mathcal{N}$ of $\Y$ contained in $\mathcal{M}$ and  stable under pullbacks, pushout, containing all isomorphism and closed under composition and decomposition.
	\end{enumerate} 
\end{theorem}

@@

In this section we introduce the notion of \emph{$\X$-hypergraph}. In order to do it is necessary to have a monad $(-)^\star$ playing a role analogous to the usua \emph{Kleene star}, also known as \emph{list monad}, $(-)^\star\colon \Set\to \Set$ sending a set to the free monoid on it \cite{sakarovitch2009elements,wadler1995monads}. In\Cref{app:mon} the interested reader can find a brief review of the construction of such a monad.

\begin{convention}
	In this section we will operate inside a countably distributive and extensive cartesian category $(\X, \times, I)$ with finite limits. As shown in \Cref{app:mon}, then $\mo{U}\colon \mono{\X}\to \X$ has a left adjoint $\mo{F}$ and $\mo{F}\circ \mo{U}$ preserves pullbacks. Denoting by $(-)^\star$ such composition, we can assume that $X^\star$ is the coproduct $\sum_{n\in \mathbb{N}}X^n$.
\end{convention}

@@

\subsection{The category of $\X$-hypergraphs}

We will start this section with the definition of $\X$-hypergraphs and we will see how to label them with an algebraic signature.

\begin{definition}An \emph{$\X$-hypergraph} is a 4-uple $\mathcal{G}:=(E_\mathcal{G}, V_\mathcal{G}, s_\mathcal{G}, t_\mathcal{G})$ made by two objects $E_\mathcal{G}$ and $V_\mathcal{G}$, called respectively the object of \emph{hyperedges} and \emph{nodes}, plus a pair of \emph{source} and \emph{target} arrows  $s_\mathcal{G}, t_\mathcal{G}\colon E_\mathcal{G}\rightrightarrows V_\mathcal{G}^\star$. A \emph{hypergraph morphism} $(E_\mathcal{G}, V_\mathcal{G}, s_\mathcal{G}, t_\mathcal{G})\to (E_\mathcal{H}, V_\mathcal{H}, s_\mathcal{H}, t_\mathcal{H})$ is a pair $(h,k)$ of morphisms $h\colon E_\mathcal{G}\to E_\mathcal{H}$, $k\colon V_\mathcal{G}\to V_\mathcal{H}$ such that the following diagrams commute.

	\[\xymatrix{ E_{\mathcal{G}} \ar[d]_{h} \ar[r]^{s_{\mathcal{G}}}& V^\star_{\mathcal{G}}  \ar[d]^{k^\star}& E_{\mathcal{G}} \ar[r]^{t_{\mathcal{G}}} \ar[d]_{h} & V^\star_{\mathcal{G}} \ar[d]^{k^\star}  \\ E_{\mathcal{G}} \ar[r]_{s_{\mathcal{H}}} & V^\star_{\mathcal{H}} & E_{\mathcal{G}} \ar[r]_{t_{\mathcal{H}}} & V^\star_{\mathcal{H}} }\]
	We define $\hyp_\X$ to be the resulting category.
\end{definition}

Let $\pro^\star$ be the functor sending $X$ to $X^\star\times X^\star$, then we can present $\hyp_\X$ as a comma category.
\begin{proposition}\label{prop:com}
	$\hyp_\X$ is isomorphic to $\comma{\id{\X}}{\pro^\star}$
\end{proposition}
\begin{proof}
	 Define two functors $F\colon  \hyp \to \comma{\id{\Set}}{\pro^\star} $
	and $G\colon \comma{\id{\Set}}{\pro^\star}\to \hyp$	 as follows 
	\begin{align*}
		\functor[l]{(E_\mathcal{{G}}, V_\mathcal{{G}}, s_\mathcal{{G}}, t_\mathcal{{G}})}{(f,g)}{\left(E_\mathcal{{H}}, V_\mathcal{{H}}, s_\mathcal{{H}}, t_\mathcal{{H}}\right)}
		\functormapsto
		\rfunctor{\left(E_\mathcal{{G}}, V_\mathcal{{G}}, 	\left(s_\mathcal{{G}}, t_\mathcal{{G}}\right)\right)}{(f, g) }{\left(E_\mathcal{{H}}, V_\mathcal{{H}}, \left(s_\mathcal{{H}}, t_\mathcal{{H}}\right)\right)}
	\end{align*}
	 \begin{align*}
		\functor[l]{(E_\mathcal{{G}}, V_\mathcal{{G}}, p_{\mathcal{G}})}{(f,g)}{(E_\mathcal{{H}}, V_\mathcal{{H}}, p_{\mathcal{H}})}
		\functormapsto
		\rfunctor{(E_\mathcal{{G}}, V_\mathcal{{G}}, 	\pi_1\circ  p_{\mathcal{G}}, \pi_2\circ  p_{\mathcal{G}})}{(f, g) }{(E_\mathcal{{H}}, V_\mathcal{{H}}, 	\pi_1\circ  p_{\mathcal{H}}, \pi_2\circ  p_{\mathcal{H}})}
	\end{align*} 
	Now it is immediate to notice that they are one the inverse of the other.
\end{proof}

\begin{corollary}\label{prop:hypadh}
If $\X$ is (strictly) $\mathcal{M}$-adhesive	then $\hyp$ is an (strict) $\comma{\mathcal{M}}{\mathcal{M}}$-adhesive category.
\end{corollary}
\begin{proof}
	By hypothesis $(-)^\star$ preserves pullbacks, while $\pro$ is continuous by definition, thus the thesis follows from  \Cref{thm:slice-functors} and \Cref{prop:com}.
\end{proof}

@@

\begin{proposition}\label{cor:left} If $\X$ has an initial object, then the forgetful functor $U_{\X}$ which sends an hypergraph $\mathcal{G}$ to its object of nodes has a left adjoint $\Delta_{\X}$.
\end{proposition}

\begin{example}Since the initial object of $\catname{Set}$ is the empty set,  $\Delta_{\Set}(X)$ is the hypergraph which has $X$ as set of nodes and $\emptyset$ as set of hyperedges and $?_X$ as both source and target function.
\end{example}

\todo{forse possiamo anche cancellare tutti gli esempi sotto}

$\Set$-hypergraphs,  can be represented graphically. We will use dots to denote nodes and squares to denote hyperedges, the name of a node or of an hyperedge will be put near the corresponding dot or square. Sources and targets are represented by lines between dots and squares: the lines from the sources of an hyperedge will have an arrowhead in the middle pointing towards the hyperedge, while the lines to the targets will have arrowheads pointing to the target nodes.  We will decorate the arrow corresponding to the $i^{th}$ letter (i.e.~its value at $i-1$) of a target or a source with a label $i$.

@@

\subsubsection{$\hyp_\X$ as a category of functors}

Following \cite{bonchi2022string}, we can present $\hyp_\X$ as a category of functor over a suitable category.

@@

Given $F\colon \catname{H}\to \catname{Set}$, since $\X$ has countable coproducts we can define
\[E_F:=\sum_{k,l\in \mathbb{N}}F(k,l)\]

Now, for every $k$, $l$, $i$ and $j$ in $\mathbb{N}$ with $i< k$ and $j< l$ we can define $s^F_{k,l}\colon F(k,l)\to F(\bullet)^k$,  $t^F_{k,l}\colon F(k,l)\to F(\bullet)^l$ as the unique ones fitting in the diagrams below, where the vertical arrows are the projections.
\[\xymatrix{F(k,l)  \ar@{.>}[r]^{s^F_{k,l}} \ar[dr]_{F(f_i)}& F(\bullet)^{k} \ar[d]^{\pi^F_{i}} & F(k,l) \ar@{.>}[r]^{t^F_{k,l}} \ar[dr]_{F(f_{k+j})} & F(\bullet)^{l} \ar[d]^{\pi^F_{j}} \\ & F(\bullet) && F(\bullet)}\]

In turn, these arrows allow us to consider
$s_F, t_F\colon E_F\rightrightarrows F(\bullet)^{\star}$ as the unique arrows fitting in the diagrams below, where the vertical arrows are coprojections.
\[\xymatrix{F(k,l) \ar[d]_{a^F_{k,l}}  \ar[r]^{s^{F}_{k,l}}& F(\bullet)^{k} \ar[d]^{b^F_{k}} & F(k,l) \ar[d]_{a^F_{k,l}}  \ar[r]^{t^{F}_{k,l}}& F(\bullet)^{l} \ar[d]^{b^F_{l}}\\ E_F \ar@{.>}[r]_-{s_F}& F(\bullet)^\star & E_F \ar@{.>}[r]_-{t_F}& F(\bullet)^\star}\]

Let $\mathcal{G}_F$ be the resulting hypergraph. Now, every $\eta\colon F\rightarrow H$ in $\catname{Set}^{\catname{H}}$ has components $\eta_{k,l}\colon F(k,l)\to H(k,l)$, $\eta_{\bullet}\colon F(\bullet)\to H(\bullet)$. In particular we can define $\hat{\eta}\colon E_F\rightarrow E_H$ as $\sum_{k,l\in \mathbb{N}}\eta_{k,l}$. Now, on the one hand, we have
\[\begin{split}
	\pi^H_i\circ \eta^k_\bullet \circ  s^F_{k,l}&=\eta_\bullet \circ \pi^F_i \circ s^F_{k,l}\\&=\eta_\bullet \circ F(f_i)\\&=H(f_i)\circ \eta_{k,l}\\&=\pi^H_i\circ s^{H}_{k,l}\circ \eta_{k,l}
\end{split} \qquad \begin{split}
\pi^H_j\circ \eta^k_\bullet \circ  t^F_{k,l}&=\eta_\bullet \circ \pi^F_j \circ t^F_{k,l}\\&=\eta_\bullet \circ F(f_{k+j})\\&=H(f_{k+j})\circ \eta_{k,l}\\&=\pi^H_j\circ t^{H}_{k,l}\circ \eta_{k,l}
\end{split} \]
so that $\eta^k_\bullet \circ  s^F_{k,l}=s^{H}_{k,l}\circ \eta_{k,l}$ and $eta^k_\bullet \circ  t^F_{k,l}= t^{H}_{k,l}\circ \eta_{k,l}$. On the other hand, these equalities give us the following:
\[\begin{split}
	\eta_{\bullet}^\star \circ s_{F}\circ a^F_{k,l}&=\eta_{\bullet}^\star \circ b^F_k\circ s^F_{k,l}\\&=b^H_k\circ \eta^k_{\bullet}\circ s^{F}_{k,l} \\&=b^H_{k}\circ s^H_{k,l}\circ \eta_{k,l}\\&=s_H\circ a^H_{k,l}\circ \eta_{k,l}\\&=	s_H\circ \hat{\eta}\circ a^F_{k,l}
\end{split} \qquad \begin{split}
	\eta_{\bullet}^\star \circ s_{F}\circ a^F_{k,l}&=\eta_{\bullet}^\star \circ b^F_k\circ t^F_{k,l}\\&=b^H_k\circ \eta^k_{\bullet}\circ t^{F}_{k,l} \\&=b^H_{k}\circ t^H_{k,l}\circ \eta_{k,l}\\&=t_H\circ a^H_{k,l}\circ \eta_{k,l}\\&=	t_H\circ \hat{\eta}\circ a^F_{k,l}
\end{split}\]
which, in turn, entail that  that the following squares commute
\[\xymatrix{E_F \ar[r]^{s_F} \ar@{.>}[d]_{\hat{\eta}}& F(\bullet)^\star \ar[d]^{\eta^\star_\bullet} & E_F \ar[r]^{t_F} \ar[d]_{\hat{\eta}} & F(\bullet)^\star \ar[d]^{\eta^\star_\bullet}\\ E_H \ar[r]_{s_H} & H(\bullet)^\star & E_H \ar[r]_{t_H}& H(\bullet)^\star}\]

This is the same as saying that $\eta$ induces a morphism $(\hat{\eta}, \eta_{\bullet})\colon \mathcal{G}_F\to \mathcal{G}_H$. It is now clear that sending $F$ to $\mathcal{G}_F$ and $\eta$ to $(\hat{\eta}, \eta_{\bullet})$ defines a faithful functor $\mathcal{G}_{-}\colon \catname{\X}^{\catname{H}}\to \hyp_\X$.

\begin{remark}\label{rem:utile}
For every $k\in \mathbb{N}$ consider the following pullback diagram	
	\[\xymatrix{F(\bullet)_k  \ar[r]^{q_1} \ar[d]_{!_{F(\bullet)_k}}& F(\bullet)^\star \ar[d]^{\lgh_{F(\bullet)}} \\ 1 \ar[r]_{o_k} & 1^\star}\]
	Now, notice that 
	\[\begin{split}
		\lgh_{F(\bullet)}\circ b^{F}_k\circ s^{F}_{k,l}&=o_{k}\circ !_{F(\bullet)^k}\circ s^F_{k,l}\\&=o_k\circ !_{F(k,l)}
	\end{split} \qquad \begin{split}
	\lgh_{F(\bullet)}\circ b^{F}_k\circ t^{F}_{k,l}&=o_{k}\circ !_{F(\bullet)^k}\circ t^F_{k,l}\\&=o_k\circ !_{F(k,l)}
	\end{split}\]
so that we have two diagrams:
	\[\xymatrix{F(k,l)\ar[r]^{s^F_{k,l}} \ar@{.>}[dr]^{u^1_{k,l}}  \ar@/_.5cm/[ddr]_{!_{F(k,l)}}&F(\bullet)^k \ar[dr]^{b^F_k}&& F(k,l) \ar@/_.5cm/[ddr]_{!_{F(k,l)}} \ar[r]^{s^F_{k,l}} \ar@{.>}[dr]^{u^2_{k,l}} &F(\bullet)^k \ar[dr]^{b^F_k}\\ & F(\bullet)_k  \ar[r]^{q_{k}} \ar[d]_{!_{F(\bullet)_k}}& F(\bullet)^\star \ar[d]^{\lgh_{F(\bullet)}} && F(\bullet)_k  \ar[r]^{q_{k} }\ar[d]_{!_{F(\bullet)_k}}& F(\bullet)^\star \ar[d]^{\lgh_{F(\bullet)}} \\ &  1 \ar[r]_{o_k} & 1^\star && 1 \ar[r]_{o_k} & 1^\star}\]

	Moreover we have
	\[\begin{split}
		s_F\circ a^F_{k,l}&=b^F_k\circ s^{F}_{k,l}\\&=
		q_{k}\circ u^1_{k,l}	\end{split} \qquad \begin{split}
		t_F\circ a^F_{k,l}&=b^F_k\circ t^{F}_{k,l}\\&=
		q_{k}\circ u^2_{k,l}	\end{split}  \]

	Therefore these two arrows, in turn, yield the dotted $v^1_{k,l}\colon F(k,l)\to F(\bullet)_k $ and $v^2_{k,l}\colon F(k,l)\to F(\bullet)_l$ as in the diagram below, where the right and bottom squares are pullbacks.
	\[\xymatrix@C=45pt{F(k,l)  \ar[dr]^{a^F_{k,l}} \ar@{.>}[r]^{v^2_{k,l}} \ar@{.>}[d]^{v^1_{k,l}} \ar@/^.8cm/[rr]^{u^2_{k,l}} \ar@/_.5cm/[dd]_{u^1_{k,l}}& P_{l} \ar[r]_{p_{l,1}} \ar[d]_{p_{l,2}}& F(\bullet)_{l} \ar[d]^{q_{l}}\\ P_k \ar[r]^{p_{k,2}} \ar[d]^{p_{k,1}}& E_F \ar[r]_{s_F}  \ar[d]^{t_F}& F(\bullet)^\star \\ F(\bullet)_k  \ar[r]_{q_{k}}& F(\bullet)^\star}\]

 By hypothesis, $\X$ is countably extensive, thus for every $k, l\in \mathbb{N}$, the following diagrams are pullbacks.
 \[\xymatrix{F(k,l) \ar[r]^{a^F_{k,l}} & E_F \ar[d]^{}\\ 
 F(\bullet)^k& F(\bullet)^\star}\]

 Let now $d_1\colon Z\to P_l$ and $d_2\colon Z\to P_k $ be two arrows such that
 \[p_{k,2}\circ d_2=p_{t,2}\circ d_1\]

 Then we have
 \begin{align*}
 t_{F}\circ 	p_{t_2}\circ d_1&= \end{align*}

 Showing, finally, that 
\end{remark}

\begin{proposition}
	$\hyp_\X$ is equivalent to the category $\catname{\X}^{\catname{H}}$.
\end{proposition}
\begin{proof}
